---
title: "AT-HOME PD Data Pre/Post State of Emergency"
output:
  html_document:
    df_print: paged
---


```{r include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=4) 
```

```{r include=FALSE}
library(synapser)
library(tidyverse)
library(lubridate)

synLogin()

MERGED_DATASET <- "syn18693245"
STATE_OF_EMERGENCY_DATE <- lubridate::as_date("2020-03-13")
WINDOW_RADIUS <- lubridate::days(120)
```

**A dash-dot line on plots corresponds to `r STATE_OF_EMERGENCY_DATE`, the assumed state of emergency (SoE) date.**

```{r include=FALSE}
mpower_surveys <- c("Background-v1", "Demographics-v1", "Engagement-v1",
                    "Medication-v1", "Medication-v3", "Motivation-v1",
                    "Symptoms-v3", "Triggers-v3")
mpower_active_tasks <- c("Tremor-v3", "Tapping-v4", "WalkAndBalance-v1")
passive_activities <- c("PassiveGait-v1", "PassiveDisplacement-v4")

get_ahpd_guids <- function(clinical = NULL, include_drop_outs = T) {
  if (is.null(clinical)) {
    clinical <- synhelper::synGetFile("syn17051543") # Rochester clinical
  }
  drop_outs <- clinical %>% 
    filter(subj_status == paste("Subject discontinued participation before",
                                "the planned study conclusion")) %>% 
    distinct(guid)
  guids <- clinical %>%
    filter(redcap_event_name == "Baseline (Arm 1: Arm 1)",
           !str_detect(guid, "TEST"),
           viscompltyn == "Yes") %>% 
    distinct(guid) %>% 
    drop_na()
  if (!include_drop_outs) {
    guids <- guids %>% 
      anti_join(drop_outs)
  }
  return(guids)
}
```

```{r include=FALSE}
get_merged_dataset <- function() {
  ahpd_guids <- get_ahpd_guids()
  window_of_interest <- c(STATE_OF_EMERGENCY_DATE - WINDOW_RADIUS,
                          STATE_OF_EMERGENCY_DATE + WINDOW_RADIUS)
  merged_dataset <- synTableQuery(
    glue::glue("SELECT * FROM {MERGED_DATASET}"))$asDataFrame() %>% 
    as_tibble() %>% 
    inner_join(ahpd_guids, by = "guid") %>% 
    drop_na(createdOn) %>% 
    filter(createdOn > window_of_interest[[1]],
           createdOn < window_of_interest[[2]]) %>% 
    mutate(createdOnDate = lubridate::as_date(createdOn))
  return(merged_dataset) 
}

merged_dataset <- get_merged_dataset()
```

## mPower

Plots of mPower data show a 180-day window centered on the SoE date. That is, all participants had a single study burst window within the 90-day window shown before the SoE and all participants had a single study burst window in the 90-day interval shown after the SoE.

### Passive Data Contributions

```{r echo=FALSE}
plot_passive_records <- function(merged_dataset) {
  # We only want to consider a 90-day window on each side of the SoE
  window_of_interest <- c(STATE_OF_EMERGENCY_DATE - lubridate::days(90),
                          STATE_OF_EMERGENCY_DATE + lubridate::days(90))
  merged_dataset <- merged_dataset %>% 
    filter(createdOn > window_of_interest[[1]],
           createdOn < window_of_interest[[2]])
  merged_dataset %>%
    filter(source == "MPOWER",
           activity %in% passive_activities) %>%
    ggplot(aes(createdOnDate)) +
    geom_bar(aes(fill = activity)) +
    geom_vline(xintercept = STATE_OF_EMERGENCY_DATE, linetype="twodash") +
    theme_minimal() +
    labs(title = "Passive Records", x = "Date", y = "Count")
}

plot_passive_records(merged_dataset)
```

```{r, include=FALSE}
SIGNIFICANT_AMOUNT_OF_RECORDS <- 10
```

The next plot attempts to show how much of this data is useable. Records contributed by users who completed a given activity `r SIGNIFICANT_AMOUNT_OF_RECORDS` or more times *both* before and after SoE are colored whereas other records are grayed out.
```{r echo=FALSE}
plot_pre_post_passive <- function(merged_dataset) {
  # We only want to consider a 90-day window on each side of the SoE
  window_of_interest <- c(STATE_OF_EMERGENCY_DATE - lubridate::days(90),
                          STATE_OF_EMERGENCY_DATE + lubridate::days(90))
  merged_dataset <- merged_dataset %>% 
    filter(createdOn > window_of_interest[[1]],
           createdOn < window_of_interest[[2]])
  pre_emergency <- merged_dataset %>%
    filter(source == "MPOWER",
           activity %in% passive_activities,
           createdOnDate <= STATE_OF_EMERGENCY_DATE)
  post_emergency <- merged_dataset %>%
    filter(source == "MPOWER",
           activity %in% passive_activities,
           createdOnDate > STATE_OF_EMERGENCY_DATE)
  pre_emergency_significant_contributors <- pre_emergency %>%
    group_by(guid) %>% 
    count(activity) %>% 
    filter(n >= SIGNIFICANT_AMOUNT_OF_RECORDS)
  post_emergency_significant_contributors <- post_emergency %>% 
    group_by(guid) %>% 
    count(activity) %>% 
    filter(n >= SIGNIFICANT_AMOUNT_OF_RECORDS)
  significant_contributors <- post_emergency_significant_contributors %>% 
    inner_join(select(pre_emergency_significant_contributors, -n),
               by = c("guid", "activity")) %>% 
    distinct(guid, activity)
  number_of_significant_contributors <- n_distinct(significant_contributors$guid)
  records_from_significant_contributors <- merged_dataset %>% 
    inner_join(significant_contributors,
               by = c("guid", "activity"))
  number_of_significant_records <- n_distinct(records_from_significant_contributors$recordId)
  date_counts <- merged_dataset %>% 
    count(createdOnDate)
  annotation_y_location <- max(date_counts$n)
  annotation_x_location <- max(date_counts$createdOnDate) - lubridate::days(15)
  merged_dataset %>%
    filter(source == "MPOWER",
           activity %in% passive_activities) %>%
    ggplot(aes(createdOnDate)) +
    geom_bar(color = "lightgray", alpha = 0.6) +
    geom_bar(data = records_from_significant_contributors,
             mapping = aes(createdOnDate, fill = activity)) +
    geom_vline(xintercept = STATE_OF_EMERGENCY_DATE, linetype="twodash") +
    annotate("text", x=annotation_x_location, y = annotation_y_location,
             label = glue::glue("Significant contributors: {number_of_significant_contributors}")) +
    annotate("text", x=annotation_x_location, y = annotation_y_location - 40,
             label = glue::glue("(Significant) records: {number_of_significant_records}")) +
    theme_minimal() +
    labs(title = "Passive Records From Significant Contributors", x = "Date", y = "Count")
}

plot_pre_post_passive(merged_dataset)
```

### Active Task Contributions

Number of active tasks performed each day is **highly** correlated across tasks.

```{r echo=FALSE}
plot_active_tasks <- function(merged_dataset) {
  # We only want to consider a 90-day window on each side of the SoE
  window_of_interest <- c(STATE_OF_EMERGENCY_DATE - lubridate::days(90),
                          STATE_OF_EMERGENCY_DATE + lubridate::days(90))
  merged_dataset <- merged_dataset %>% 
    filter(createdOn > window_of_interest[[1]],
           createdOn < window_of_interest[[2]])
  merged_dataset %>%
    filter(source == "MPOWER",
           activity %in% mpower_active_tasks) %>%
    ggplot(aes(createdOnDate)) +
    geom_bar(aes(fill = activity)) +
    geom_vline(xintercept = STATE_OF_EMERGENCY_DATE, linetype="twodash") +
    theme_minimal() +
    labs(title = "Active Task Records", x = "Date", y = "Count")
}

plot_active_tasks(merged_dataset)
```

```{r echo=FALSE}
plot_pre_post_active <- function(merged_dataset) {
  window_of_interest <- c(STATE_OF_EMERGENCY_DATE - lubridate::days(90),
                          STATE_OF_EMERGENCY_DATE + lubridate::days(90))
  merged_dataset <- merged_dataset %>% 
    filter(createdOn > window_of_interest[[1]],
           createdOn < window_of_interest[[2]])
  pre_emergency <- merged_dataset %>%
    filter(source == "MPOWER",
           activity %in% mpower_active_tasks,
           createdOnDate <= STATE_OF_EMERGENCY_DATE)
  post_emergency <- merged_dataset %>%
    filter(source == "MPOWER",
           activity %in% mpower_active_tasks,
           createdOnDate > STATE_OF_EMERGENCY_DATE)
  pre_emergency_significant_contributors <- pre_emergency %>%
    group_by(guid) %>% 
    count(activity) %>% 
    filter(n >= SIGNIFICANT_AMOUNT_OF_RECORDS)
  post_emergency_significant_contributors <- post_emergency %>% 
    group_by(guid) %>% 
    count(activity) %>% 
    filter(n >= SIGNIFICANT_AMOUNT_OF_RECORDS)
  significant_contributors <- post_emergency_significant_contributors %>% 
    inner_join(select(pre_emergency_significant_contributors, -n),
               by = c("guid", "activity")) %>% 
    distinct(guid, activity)
  number_of_significant_contributors <- n_distinct(significant_contributors$guid)
  records_from_significant_contributors <- merged_dataset %>% 
    inner_join(significant_contributors,
               by = c("guid", "activity"))
  number_of_significant_records <- n_distinct(records_from_significant_contributors$recordId)
  date_counts <- records_from_significant_contributors %>% 
    count(createdOnDate)
  #annotation_y_location <- max(date_counts$n)
  annotation_x_location <- max(date_counts$createdOnDate) - lubridate::days(15)
  merged_dataset %>%
    filter(source == "MPOWER",
           activity %in% mpower_active_tasks) %>%
    ggplot(aes(createdOnDate)) +
    geom_bar(color = "lightgray", alpha = 0.6) +
    geom_bar(data = records_from_significant_contributors,
             mapping = aes(createdOnDate, fill = activity)) +
    geom_vline(xintercept = STATE_OF_EMERGENCY_DATE, linetype="twodash") +
    annotate("text", x=annotation_x_location, y =100,
             label = glue::glue("Significant contributors: {number_of_significant_contributors}")) +
    annotate("text", x=annotation_x_location, y = 90,
             label = glue::glue("(Significant) records: {number_of_significant_records}")) +
    theme_minimal() +
    labs(title = "Active Tasks From Significant Contributors", x = "Date", y = "Count")
}

plot_pre_post_active(merged_dataset)
```

## Clinical (Rochester/CTCC)

Clinical tele-visits only happen once a year.

```{r echo=FALSE}
plot_clinical <- function(merged_dataset) {
  merged_dataset %>% 
    filter(source == "ROCHESTER") %>% 
    distinct(guid, createdOnDate) %>% 
    ggplot(aes(createdOnDate)) +
    geom_histogram(binwidth = 7) +
    geom_vline(xintercept = STATE_OF_EMERGENCY_DATE, linetype="twodash") +
    theme_minimal() +
    labs(title = "Clinical tele-visits each week",
         x = "Date", y = "Visits This Week")
}

plot_clinical(merged_dataset)
```

## Fox Insight Surveys

Fox Insight surveys are taken at regular intervals which depend on the survey. The most frequent surveys are meant to be taken quarterly (relative to registration date). Participants are given a 30-day window to complete the surveys, hence we have plotted all data over a 240-day window centered on the SoE date. (The largest possible interval between quarterly surveys is 90+30=120 days, hence we need to consider the 120 days before the SoE).

The number of surveys taken after the SoE which were also completed by the same participant in the time window shown before the SoE are plotted in red.

```{r echo=FALSE}
plot_fox <- function(merged_dataset) {
  pre_emergency_surveys <- merged_dataset %>% 
    filter(source == "MJFF",
           createdOnDate <= STATE_OF_EMERGENCY_DATE) %>% 
    distinct(guid, createdOnDate, activity) %>% 
    mutate(completed_before_emergency = TRUE)
  post_emergency_surveys <- merged_dataset %>% 
    filter(source == "MJFF",
           createdOnDate > STATE_OF_EMERGENCY_DATE) %>% 
    distinct(guid, createdOnDate, activity) %>% 
    full_join(select(pre_emergency_surveys, -createdOnDate),
              by = c("guid", "activity")) %>% 
    mutate(completed_before_emergency = replace_na(
      completed_before_emergency, FALSE))
  # these surveys are "paired" with the same survey this user completed
  # before a state of emergency was declared
  post_emergency_surveys_paired <- post_emergency_surveys %>% 
    filter(completed_before_emergency)
  merged_dataset %>%
    filter(source == "MJFF") %>% 
    ggplot(aes(createdOnDate)) +
    geom_histogram(binwidth = 7) +
    geom_histogram(data = post_emergency_surveys_paired,
                   binwidth = 7, fill = "red") +
    geom_vline(xintercept = STATE_OF_EMERGENCY_DATE, linetype="twodash") +
    theme_minimal() +
    labs(title = "Fox Insight Forms Completed",
         x = "Date", y = "Forms Completed This Week")
}

plot_fox(merged_dataset)
```
